### 2018-04-19 statistical-quorum.tex
The vote -> voting.

\documentclass[format=acmsmall, review=true, screen=true, anonymous=true]{acmart}

\usepackage[english]{babel}

\usepackage{booktabs} % For formal tables

\usepackage[ruled]{algorithm2e} % For algorithms
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

\usepackage{todonotes}
\usepackage{bookmark}
\usepackage[T1]{fontenc}
\usepackage{microtype}
\usepackage[utf8]{inputenc}

% Metadata Information
%\acmJournal{TWEB}
%\acmVolume{9}
%\acmNumber{4}
%\acmArticle{39}
%\acmYear{2010}
%\acmMonth{3}
%\copyrightyear{2009}
%\acmArticleSeq{9}

% Copyright
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
%\acmDOI{0000001.0000001}

% Paper history
%\received{February 2007}
%\received[revised]{March 2009}
%\received[accepted]{June 2009}

% Document starts
\begin{document}
% Title portion. Note the short title for running heads
\title[Improving credibility and efficiency of decision-making]{Improving credibility and efficiency of decision-making \\ within democratic communities via statistical quorum}

\author{Mitar Milutinovic}
\affiliation{%
  \institution{UC Berkeley}}
\email{mitar@cs.berkeley.edu}
\author{Maxinder S. Kanwal}
\affiliation{%
  \institution{UC Berkeley}}
\email{maxowaxo2013@gmail.com}
\author{David Marn}
\affiliation{%
  \institution{UC Berkeley}}
\email{david.marn@gmail.com}

\begin{abstract}
The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community.
While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally
different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to
maximize productivity?
We propose a novel \emph{statistical quorum}, derived using first principles, as one metric the community can use to
focus attention, in turn making the entire decision-making process dynamic.
We developed a web application centered around this metric and tested it in a 140-person community that holds weekly
councils over the course of 16 weeks.
We found that the community was able to use the web application to simultaneously improve the credibility and efficiency
of their decision-making process.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10003120.10003130.10003131.10003570</concept_id>
<concept_desc>Human-centered computing~Computer supported cooperative work</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003120.10003130.10011764</concept_id>
<concept_desc>Human-centered computing~Collaborative and social computing devices</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10010405.10010476.10010936.10003590</concept_id>
<concept_desc>Applied computing~Voting / election technologies</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Human-centered computing~Computer supported cooperative work}
\ccsdesc[500]{Human-centered computing~Collaborative and social computing devices}
\ccsdesc[500]{Applied computing~Voting / election technologies}

%
% End generated code
%

\keywords{Decision making; voting; democracy; quorum; credibility; statistical quorum}

\maketitle

% The default list of authors is too long for headers.
\renewcommand{\shortauthors}{M. Milutinovic et al.}

\section{Introduction}

In many historical and recent elections it is not infrequent that we get close results (see Table~\ref{tab:example}).
Especially when this is combined with low voter turnout, how can we have confidence in results?
What if elections would be a day longer? What if they would be on a different day? What if
more people would participate? Would results still be the same?

\begin{table}
  \centering
  \begin{small}
  \begin{tabular}{l r r r r r r r}
    {\small\textit{Vote}} & \multicolumn{2}{r}{\small \textit{First option}} & \multicolumn{2}{r}{\small \textit{Second option}} & {\small \textit{Total eligible}} & {\small \textit{Turnout}} & {\small \textit{Decision}} \\
    & & & & & {\small \textit{voters}} & & {\small \textit{credibility}} \\
    \midrule
    Brexit & 17,410,742 & $51.9\%$ & 16,141,241 & $48.1\%$ & 46,500,001 & $72.2\%$ & $59.8\%$ \\
    Presidential elections & 62,984,825 & $48.9\%$ & 65,853,516 & $51.1\%$ & 230,585,915 & $55.9\%$ & $52.6\%$ \\
  \end{tabular}
  \end{small}
  \bigskip\caption{Votes cast in Brexit, a 2016 referendum for United Kingdom's withdrawal from the European Union~\protect\cite{brexitresults},
  and 2016 USA presidential elections, popular vote~\protect\cite{usaresults1,usaresults2}. Simplified to only two options and without
  invalid votes. The last column shows computed decision credibility as described in Section~\ref{sec:decision-credibility}. We see that
  credibility is low, which we can mostly attribute to the fact that there is small difference in votes between both options so
  with more votes any option could end up winning.}\label{tab:example}
  \vspace{-0.3cm}
\end{table}

An underlying question is what should we do when not everyone in the population votes.

Traditionally, we can have a stance that everyone had an opportunity to vote and express
their preference, so whatever are results based on those who voted are legitimate results.
Research shows, though, that there are many reasons why people do not vote and that there
are many issues related to how elections are organized~\cite{CANCELA2016264, GIMPEL2003471}.

Alternatively, we can see voting as a survey and cast votes as a sample, although imperfect
and non-randomly selected, from a larger population, with a goal to infer the preference
of the whole population. In this setting we can use statistical methods to evaluate the
confidence we can have in the results.

%Moreover, we can see overall participation by everyone in the population as a
%limited resource, a collective attention, we have at our disposal to make collective decisions.
%Traditionally, a common solution to the low voter turnout is to focus on
%increasing the turnout and through that increasing the collective attention available
%for decision making.
%But increasing collective attention for decision making is at odds with other
%demands for the collective and individual attention.
%But once we have a statistical measure of results, we can utilize it to
%optimize the allocation of the collective attention by directing participation towards
%contentious issues.

Our contributions are:

\begin{itemize}
\item We developed a framework of computing statistical credibility of results
based on sample size, votes cast, and voting duration.
\item Based on this framework we developed a novel concept of \emph{statistical quorum}
as a threshold of votes required to deem results credible and legitimate.
\item We implemented web application \emph{PeerMind},
integrating credibility of results and statistical quorum, and designed
a workflow of its use in a decision making processes.
\item We deployed the application in a democratically run community of 140
members for 16 weeks.
\item Using the analysis of decision making process before and after introduction
of the application, using surveys, and using semi-structured interviews, we show
how use of statistical quorum increases credibility of decisions,
improves utilization of collective attention, increases and diversifies overall participation,
and increases number of decisions made.
\end{itemize}

The rest of the paper is structured as follows. In Section~\ref{sec:related-work} we
present related work. We continue by describing a framework of computing statistical
credibility of results and statistical quorum in Section~\ref{sec:decision-credibility}.
We present our methodology in Section~\ref{sec:methodology} after which we do qualitative and
quantitative evaluation in Section~\ref{sec:evaluation}. We discuss results in
Section~\ref{sec:discussion} and conclude in Section~\ref{sec:conclusion}. At the end we
have Appendix where we derive formulas presented in Section~\ref{sec:decision-credibility}.

\section{Related Work}
\label{sec:related-work}

%\todo[inline]{Threshold and quorum at \url{https://www.publications.parliament.uk/pa/ld200910/ldselect/ldconst/99/99.pdf}.
%How do some claim that quorum are in fact bad, and counter democratic, but do not offer a solution how to evaluate decisions.}
%
%\todo[inline]{Voting auditing work.}
%
%\todo[inline]{Delegation as way to increase legitimacy, Liquid Democracy, etc.}

\section{Decision Credibility and Statistical Quorum}
\label{sec:decision-credibility}

%Range voting (also known as score voting) is voting where members of a community express their positions on open proposals by assigning
%each a score on a fixed scale, in the context of this paper, between $-1$ and $+1$.
%A proposal passes if the average score of all votes on a position is positive.
%When multiple competing proposals are being voted on, the one with highest average score wins.
%
%Prior work has shown that this approach generalizes many previous approaches to voting and allows members to express their
%positions in a highly informative and precise manner.
%We find it especially suitable for a web application because many web users are used to use it when ranking products
%and giving other ``five-star'' feedback.
%Moreover, if later on a new proposal is added to a set of existing proposals on which voting was already open, only the
%new proposal needs to be evaluated by members, as opposed to alternative voting systems where the relative rankings of
%each proposal would need reassessing (assuming members use a consistent personal scale when voting).

A statistical quorum uses statistical hypothesis testing to infer the credibility of the outcome
of an ongoing voting before the community adopts the outcome as being truly representative of all
members' positions.
For example, in a community of 100 members, if 35 members support a specific proposal and 0 oppose,
then we can conclude with high probability that, had all 100 members taken the time to vote, a
majority of them would be in support of the proposal.
Hence, despite the reality of 65 members not having voted, the outcome of the 35-member vote is
credible enough to be taken as representative of the entire community's desire.
On the other hand, if 80 members vote, where 41 support the proposal and 39 oppose, then we
cannot yet conclude that the decision to accept the proposal is credible enough for the entire
community to officially adopt, despite the significant participation in voting.
Instead, the suggested course of action would be to continue gathering more votes and/or further
discussing the proposal until a credible outcome is established.

By utilizing a statistical quorum, a community can actively direct attention and participation
within the community towards contentious proposals, while avoiding wasting attention on
noncontroversial proposals.

\subsection{Background}
Consider a community that needs to collectively decide which proposal (if any) to accept out of
a set of competing proposals.
To act in democratic fashion, the community may wish to select a winning proposal that either has
the support of a \{simple, super\}-majority of all members, or instead is demonstrably preferred
over all other competing proposals.
(In some cases, the community may want a winning proposal to satisfy both criteria.)

Ideally, we do not want to poll all members in order to officially accept a proposal, but we do
want some level of certainty (e.g., $90\%$ certainty) that the population is not accidentally
accepting an undesired proposal, which in this case encompasses accepting a proposal that is
not supported by a majority of members and/or is not clearly preferred over all other competing
alternatives.

The primary motivation behind using a statistical quorum is to guarantee a level of credibility
before making a decision on behalf of the entire community.

\subsection{Problem Statement}
We wish to compute the probability that the bias in a community towards supporting a
proposal is greater than some fraction of the population (e.g., $50\%$ for a simple majority,
$67\%$ for a super majority).

In addition, we wish to compute the probability that the community bias towards supporting
a specific proposal is greater than the community bias towards supporting any competing proposals.

\subsection{Assumptions}
We assume that there is a fixed latent bias for the support of a proposal within the community.
We take each individual vote to be independent and identically distributed (i.i.d.) according
to a Bernoulli distribution around this bias.
It should be noted that the independence assumption may not hold in some settings, such as when
votes are accumulated over time (one may observe surges of correlated votes).
We will partially address this when refining our original formulation of a statistical quorum
into a \textit{dynamic} statistical quorum.

Without having collected any votes, we assume that the community bias may be any value between
$0\%$ and $100\%$ with equal probability (uniform prior).

\subsection{Voting}
We allow for range voting where each member of the community may choose any real-valued vote
$v \in [-1,+1]$ to represent their vote, with $v = -1$ representing full opposition and $v=+1$
representing full support.
In most applications, it is sensible to quantize $v$ into a fixed number of bins spanning
the full range, hence our treatment of $v$ as a discrete random variable.
Because we treat individual votes as a Bernoulli random variable, one can consider $v$ as
an individual's average vote.
In the following subsection, we rescale $v$ to be in the range $[0,1]$ via the
notation $\frac{v + 1}{2}$.

We allow voters to ``abstain'' from voting.
We treat abstentions as indicating that a voter is removing themselves from the community on a
given vote.
It is important to distinguish an ``abstention'' (removing one's self from the voting population)
from voting $0$ (indicating indifference).

\subsection{Bayesian Theory}

\todo[inline]{Weave in a single concrete example to ground all the abstract theory so that the
entire derivation and motivation has a clear and cohesive thread from start to finish.}

After collecting an initial set of votes from members, we want to use that data to construct
a prior distribution to help estimate the posterior probability that the proposal is acceptable
to the entire community.
Because the sum of the collected votes follows a binomial distribution, a commonplace and
convenient choice for the form of the prior on the community's bias is the Beta distribution
(due to its conjugacy with the binomial distribution).
Together, they form the Beta-binomial distribution (related to the P\'olya urn model).

After extracting the sufficient statistics from raw votes --- namely, the average score
and the number of abstentions --- we can adjust the effective community size and corresponding
threshold.
In addition, we can calculate the effective number of ``yes'' ($+1$) and ``no'' ($-1$) votes:

Let
\begin{description}
\item[$C = $] $\{X_1,\ldots, X_n\}$, the set of competing proposals,
\item[$N = $] population size,
\item[$X_i = $] $\{v_1, \ldots, v_{m_i}\}$, the $m_i$ raw votes for proposal $i$, after excluding
abstentions,
\item[$N_i = $] $(N - \textrm{number of abstentions on } X_i)$, the effective population voting
on proposal $X_i$,
\item[$\alpha_i = $] $\sum\limits_{v \in X_i}\left(\frac{v+1}{2}\right)$, the effective number
of ``yes'' votes on proposal $X_i$,
\item[$\beta_i = $] $m_i - \alpha_i$, the effective number of ``no'' votes on proposal $X_i$,
\item[$K_i = $] $N_i - m_i$, the number of remaining votes yet to be cast on $X_i$,
\item[$p_i = $] true latent community bias towards voting ``yes'' on $X_i$.
\end{description}

In Eqs.~\eqref{EQ2}, \eqref{EQ3}, \eqref{EQ5}, \eqref{EQ6}, \eqref{EQ7} that follow, we refer the
reader to the \hyperref[sec:derivation]{Appendix} for complete derivation details and justifications.

When considering a single proposal $X_i$ in isolation, we wish to compute:
\begin{align}\label{EQ1}
\Pr(\alpha^*_i > T_i \mid X_i),
\end{align}
where
\begin{description}
\item[$\alpha^*_i = $] the total number of ``yes'' votes for $X_i$ if we were to sample the entire
population,
\item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed
(e.g., $f=\frac{1}{2}$ for a simple majority).
\end{description}
Using a desirable prior
\begin{align}\label{EQ2}
p_i \mid X_i \sim \operatorname{Beta}(1+\frac{\alpha_i}{K_i},1+\frac{\beta_i}{K_i}),
\end{align}
corresponding to a uniform prior when no votes have been cast ($\alpha_i = \beta_i = 0$), we can
express \eqref{EQ1} as:
\begin{align}\label{EQ3}
\Pr(\alpha^*_i > T_i \mid X_i) &= 1 - {CDF}_{\nu_i}(V_i)
\end{align}
for
\begin{align*}
&\nu_i \sim \operatorname{Beta-Binomial}(1 + \frac{\alpha_i}{K_i}c, 1 + \frac{\beta_i}{K_i}c, K_i),\\
&{CDF}_{\nu_i}(V_i)  = \Pr(\nu_i \leq V_i),
\end{align*}
\todo[inline]{Note that we choose $c=10$ because it leads to intuitive results.}
where
\begin{description}
\item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes''
votes $X_i$ could receive such that it would remain below threshold.
\end{description}

When considering a proposal $X_i$ relative to other proposals in $C$, we wish to compute
\begin{align}\label{EQ4}
\Pr\left(\bigwedge_{j \neq i} (\alpha^*_i > \alpha^*_j) \mid C\right)
\end{align}

Under the assumption that the community's votes on proposals are independent across proposals
(which, in many instances, will actually give a lower bound on \eqref{EQ3} when a community's
support of one proposal is negatively correlated with its support on a competing proposal),
we can decompose \eqref{EQ3} as follows:
\begin{align}\label{EQ5}
& \Pr\left(\bigwedge_{j \neq i} (\alpha^*_i > \alpha^*_j) \mid C\right) \nonumber \\
= & \prod_{j \neq i}\left(1 - \sum_{k=\lceil\delta_{ij}\rceil}^{K_j} \Pr(\nu_j = k) \cdot {CDF}_{\nu_i}(\alpha_j + k) \right)
\end{align}
where
\begin{description}
\item[$\delta_{ij} = $] $\alpha_i - \alpha_j$, the number of effective ``yes'' votes $X_i$
leads $X_j$ by.
\end{description}

Using Eqs.~\eqref{EQ2} and \eqref{EQ4}, we can numerically compute the credibility of a
vote's outcome in two key scenarios relevant for democratic communities.
For example, a community may only consider accepting a proposal when a vote's credibility
is above $90\%$.

\subsection{Dynamic Statistical Quorum}
Oftentimes not all proposals are relevant to the entire community.
Instead, a niche proposal may only affect a subset of the community, and hence to require much
of the community to indicate their indifference and/or decide their opinion is a waste of their
attentional budget.
While the underlying formulation for a statistical quorum was to establish credibility guarantees
before making a decision, a natural consequence in common scenarios of the kind just described
leads to a reduction in decision-making efficiency.
How can we maintain credibility guarantees while simultaneously boosting efficiency?

The solution is to infer the \textit{relevant} population size that cares about a given proposal
so that we may establish credibility within only the subset population (naturally requiring less
global attention).
By replacing the absolute upper bound on this relevant population size ($N_i$) with a high
probability upper bound, we can continue to use the same formula presented in Eqs.~\eqref{EQ2}
and \eqref{EQ4}, but with more relevant parameters (namely, a smaller $N_i$, which in turn
modifies $T_i$, $K_i$, and $V_i$).

To infer $N_i$, we turn to a dynamic voting process whereby information embedded within the
dynamics is exploited.
By modeling the arrival rate of new votes as a nonhomogeneous Poisson process, we can extrapolate
into the future.
For a decaying arrival rate of new votes, a nonhomogeneous Poisson process
$\{N(t),t\geq 0\}$ probabilistically models the number of future arrivals over a given time
interval $[0,t)$ according to:
\begin{align*}
\Pr(N(t)=n)={\frac {[\Lambda (t)]^{n}}{n!}}e^{-\Lambda (t)},
\end{align*}
where
\begin{align*}
\Lambda (t)=\int _{0}^{t}\lambda (\tilde{t})\mathrm{d}\tilde{t},
\end{align*}
for some arrival rate process $\lambda (t)$.

For our purposes, we make a natural choice,
\begin{align}\label{EQ6}
\lambda (t) = \lambda_0 {(\gamma)}^{\frac{t}{K_i/N_i}},
\end{align}
where
\begin{description}
\item[$\lambda_0 =$] a time-averaged rate of vote arrivals (preferably over a unit of time that
sufficiently smooths the historical rate),
\item[$\gamma =$] a decay factor ($0 < \gamma < 1$) ideally tuned to fit real historical data
for the community via a regression analysis.
\end{description}
Now, for example, to upper bound with $95\%$ probability the remaining number of votes still
pending, we wish to solve for $\hat{K_i}$ such that the following inequality is satisfied:
\begin{align}\label{EQ7}
\lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{\Lambda^n}{n!}e^{-\Lambda} \geq 95\%,
\end{align}
where $\Lambda = \frac{\lambda_0 \cdot K_i/N_i)}{\ln{1/\gamma}}$.
Upon solving for $\hat{K_i}$, we update $N_i \leftarrow \min\{N_i,\alpha_i + \beta_i + \hat{K_i}\}$,
and can proceed to check for a statistical quorum using Eq.~\ref{EQ5}.

%\begin{figure}[ht]
%\centering
%\includegraphics[width=0.5\textwidth]{figures/}
%\caption{Simple Majority: $\alpha$ vs. $\Pr(\alpha > \frac{1}{2}N)$ for $N=100$, $\beta = 10$.}
%\label{simple_majority}
%\end{figure}

\section{Methodology}
\label{sec:methodology}

To evaluate statistical quorum in practice, we developed a web application PeerMind and
used it in a democratically run community of 140 members for 16 weeks, replacing the traditional
decision making through in-person meetings. The community is a house of 140
students living together and making collective decisions about the governance and operation of
the house, use of resources like common spaces and budgets, and changes to the house itself.

Before the introduction of the application, community had a practice of weekly meetings, councils, with all
members being invited to participate. For decision making it used variation of Robert's rules~\cite{roberts}
with majority and super-majority voting to adopt decisions, generally using hand voting and voice voting.
A fixed quorum of $30\%$ of the membership was used (42 members). Councils generally lasted between one to three hours.
Public minutes were taken.

We introduced the application in fall 2016 semester which lasted 16 weeks. We use spring 2016 semester as
a baseline for comparison, which lasted 16 weeks as well and traditional decision making was used.

To evaluate general perception of members before and after the use of the application for decision making,
we did a general survey at the end of each semester, asking members to answer general questions about the
decision making in the house that semester.

We instrumented the application to collect information on how much time users spent using the application,
measuring the time the application was focused. Other information we used for evaluation are available
through regular features of the application: how many and who made discussion items, comments, proposals
(proposed decisions on discussion items), and vote tallies per proposal.

For the baseline, we extracted similar information from council minutes: how long did the councils last, how many
discussions were opened, how many comments by which members were made, how many proposals were proposed,
and vote tallies for them.

Finally, we did semi-structured interviews after the fall semester with 7 members who were participating
both in spring and fall semesters and ask them to compare their experience between those two semesters and
evaluate the application and the use of statistical quorum.

\subsection{PeerMind}
\label{sec:peermind}

We built a full-fletched mobile-friendly web application using Meteor web framework~\cite{meteor}.
Meteor framework supports reactivity which makes the application update user interface based on actions
of other users in real time. This helps users see comments by other users as they happen, but can also see
progress during voting in real time.

Each user can open and propose new discussion items, which leads them to the main user interface for each discussion as seen in Figure~\ref{fig:peer-mind}, having three sections: commenting section on the left, summary points in the middle,
and proposed proposals with voting on the right. All content inputs fields accept rich text content in WYSWYG manner,
so users can easily add links, images, and other files.

A general workflow after creation of a discussion item is that users discuss it through comments,
which moderators summarize into summary points. At some point an user proposes a proposal which can be
further discussed through comments. Once a moderator deems that proposal is ready to be voted upon
can open it for voting. Users vote through a ballot seen in Figure~\ref{fig:ballot}.

An important design decision in the scope of this paper is that during voting, users can see the current
results of voting in real time. Results are shown on a chart depicting average score through time
as seen in Figure~\ref{fig:results}. Users can see both the current average score of all votes (score
larger than 0 means that a proposal is passing) and the credibility of current result. A community can decide
how high credibility it requires to deem the result passing statistical quorum. During our evaluation
we used 0.90 as statistical quorum.

The rationale for displaying the current results during voting is to give users feedback how voting
is progressing and allow them to engage more with the discussion. Moreover, displaying the results
also serves as additional safeguard against potential manipulations of statistical quorum through malicious
coordination.

After voting passes the statistical quorum, a moderator can close voting and a discussion item.

Web application contains other features which makes it easier to use as a decision making and discusion
platform, like notifications, and additional features which allow the application to be used also to
help run in-person meetings to discuss discussion items made
online. Facilitators of in-person meetings can use it to prepare agendas for meetings, take notes,
and display summary points of current discussion and what proposals are being voted on. Through that the
application can also serve as way to integrate both online and in-person decision making. Having all of
decision making in one platform serves as an important resource for institutional memory.

PeerMind source code is available under open source licence at \url{https://redacted.for.blind.review} and
anonymized dataset of interactions with the app, including votes, is available at \url{https://redacted.for.blind.review}.

\begin{figure}[ht]
\centering
\includegraphics[width=0.9\textwidth]{figures/peermind.png}
\caption{User interface of PeerMind web application. One can see three column structure, with commenting
section on the left, summary points in the middle, and proposed proposals with voting on the right.}
\label{fig:peer-mind}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=0.5\textwidth]{figures/ballot.png}
\caption{Ballot used for voting in PeerMind web application. Users can chose a value between -1 and 1 on a slider,
where -1 means opposition, 1 means in favor, and 0 indifference. They can also chose abstention (they
remove themselves from the population of the community for voting in question), or to not vote (default), which
means that their vote does not contribute to the quorum.}
\label{fig:ballot}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=0.6\textwidth]{figures/results.png}
\caption{Results of a vote in PeerMind web application. Results are shown as an average score through time of voting.
A user can hover over the chart to see details at that time. Users can see these results in real time during voting.}
\label{fig:results}
\end{figure}

\section{Evaluation}
\label{sec:evaluation}

We evaluate statistical quorum in two ways. First, we analyze it theoretically and explore its mathematical properties.
Then we evaluate it through real-world use of the PeerMind application by a 140 members community.

\subsection{Analysis}

\todo[inline]{Visualize the credibility and dynamics of the voting process under this framework:\\
4 Figures and 1 Table Pending}

\subsection{Use in Practice}

designed integration with system

minutes and issues with how to get data out

workflow we created -> over one in-person meetings, but because of all decision making happining online, this
didn't happen

issues with quorum
unclear how much time spend on each discussion item: how contentious the discussion really is can be seen only after voting

X people choose not to participate

\todo[inline]{Explain how we were using this for 16 weeks in fall 2016 in a community where traditionally they have
been meeting weekly and now they used it online. How X proposals have been made with Y votes.}

\todo[inline]{Explain how we compare with history: using minutes from spring 2016 for the same community. Issues with
quorum often not bein reached or being lost during the meeting. How many proposals were made then.}

\todo[inline]{Using our metric compare legitimacy of decisions being made in the spring and fall.}

\todo[inline]{We compare how much time collectively was spend on decision making based on minutes in spring 2016 and
how much based on interactions with the platform in fall 2016. Claim is that we made more decisions with less time people
having to be involved in decision making, and with better legitimacy.}

\todo[inline]{We did satisfaction surveys with decision making both in spring and fall 2016. Pull out some interesting
observation when comparing.}

\todo[inline]{Make sure that we are supporting both of our claims well: this statistical quorum is a good metric and
collective attention can be utilized better.}

\subsection{Survey}

In Figure~\ref{fig:survey-comparison} we see questions and results of the survey we made at the end of each semester.

We can see that in fall members were more satisfied with the decision making in the community and only how understandable
is the whole process received a decrease. We attribute this to the fact that the use of the web application is
a novelty and requires more effort in educating members than a traditional decision making. Moreover, decision
making in person allows one to observe others how they are participating and learn from them, instead of
being alone using the application.

\begin{figure}[ht]
\centering
\includegraphics[width=0.5\textwidth]{figures/survey-comparison.pdf}
\caption{Results of a survey at the end of spring and fall semesters about general perception of the decision making (councils) in the community.
For each question, a participant was asked to rank the answer on a scale between 1 (completely disagree) and 7 (completely agree). All members
of the community were invited to participate in the survey each time.}
\label{fig:survey-comparison}
\end{figure}

\subsection{Interviews}

\todo[inline]{We made 7 unstructured interviews. Pull all quotes explaining how they like the idea and how it seems
to work. But how there are other issues by moving decision making online: meetings also serve community building which
does not happen online. No fun. So interesting insights about using online platforms for offline communities
for making decisions which are not really answering our hypothesis, but are useful as insights we got. How some
people prefer online and offline meetings.}

\section{Discussion}
\label{sec:discussion}

\todo[inline]{Describe why is $0.90\%$ a reasonable statistical quorum and how to determine one for your community.\\
Part of the answer here is honestly just that the admin should tune the threshold (and a few of the other hyperparameters we introduce) to their intuition.  Once you set the parameters, the real benefit is that you have consistency, rather than having to always use your intuition as an admin about evaluating credibility.}

\todo[inline]{Describe interesting example how we had a bug which were showing a very low statistical quorum for a failing
vote and so the whole community very engaged for a whole month to decide and resolve it. So how they trusted the
app and the metric and how it was really able to direct the attention of the community. It seems it works even if it is
buggy or not perfect, because it has worked so well in the past that members got used to the fact that when it was saying
that there is a consensus or not on an issue that that was also their own read of the community. So for the purpose of
directing attention the metric does not have to mathematically perfect.}

\todo[inline]{Security. How to prevent a group of people to vote and by trick reaching statistical quorum.
We use dynamic statistical quorum to address this, but one other approach is also for voting to stay open and
decisions can then be in the future revoked by people adding more votes, or changing them. But for some decisions
this might not really be possible, because decisions might not be reversible. You can revoke a law, but if you built
a building as a consequence of a decision?\\
(See near/post-quorum validation phase idea in GitHub issues as one suggested solution.)}

\todo[inline]{Discuss how this might or might not lead to a more inclusive decision making: some people prefer it because
they have time to collect thoughts and can participate when they have time, some people do not because they feel
they cannot communicate as effectively, especially with lack of emotions being conveyed.}

%\section{Future Work}
%
%\todo[inline]{More work on security evaluation of potential attacks when a community is using statistical quorum to force a decision.
%Is there a game theoretical approach to it?}
%
%\todo[inline]{How to surface information about credibility to users better than just a number? Visualizations? How can we help them
%understand what exactly this is and how the math works?}

\section{Conclusion}
\label{sec:conclusion}

% Appendix
\appendix
\label{sec:derivation}

\todo[inline]{Focus on only including bare minimum math, (no ``decorative math'').  Emphasis on justifications and properties, minimal arithmetic.  Good heuristic: Every line of math should have something said about it, otherwise it is unnecessary.}

\section{Derivation of Eq.~\eqref{EQ2}}
In choosing a Beta prior, we start with $\operatorname{Beta}(1,1)$, as it represents the uniform distribution over all potential values of the community's bias.  To update the prior based on collected votes, standard theory suggests updating to $\operatorname{Beta}(1+\alpha_i,1+\beta_i)$; however, there are a few consistency issues with this update that we seek to fix:
\begin{description}
\item[$(1)$] The updated prior should not depend on the absolute number of votes $\alpha_i$, $\beta_i$, as this is not scale-invariant; rather, it should depend on the relative proportions (which are unitless instead of a number of votes),
\item[$(2)$] as $\alpha_i + \beta_i \rightarrow N_i$, $\mathbb{E}[p_i] \rightarrow \frac{\alpha_i}{\alpha_i+\beta_i}$ (at a rate independent of $N_i$),
\item[$(3)$] as $\alpha_i + \beta_i \rightarrow N_i$, $\operatorname{Var}[p_i] \rightarrow 0$ (at a rate independent of $N_i$).
\end{description}
To satisfy these consistency properties, we seek to linearly update the prior's scale parameters in a more natural way: $\operatorname{Beta}(1+\eta\alpha_i,1+\eta\beta_i)$ for some renormalizing factor $\eta$.

We can now solve for $\eta$:
\begin{align*}
\operatorname{Var}[p_i] &= \frac{(1+\eta\alpha_i)(1+\eta\beta_i)}{(2+\eta\alpha_i+\eta\beta_i)^2(3+\eta\alpha_i+\eta\beta_i)}
\end{align*}
Clearly, to satisfy $(3)$, we need $\eta \rightarrow \infty$ as $\alpha_i + \beta_i \rightarrow N_i$.  Consequently, the dominant asymptotic term becomes $\frac{1}{\eta^3 (\alpha_i + \beta_i)^3} = \frac{1}{\eta^3 N_i^3}$.  Hence, for the variance to asymptotically approach $0$ at a rate independent of $N_i$, we need $\eta \propto \frac{1}{N_i}$.

Combining these two constraints, we arrive at $\eta \propto \frac{1}{N_i}\cdot \frac{1}{K_i/N_i}$, where the latter factor leads to $\eta \rightarrow \infty$ as $\alpha_i + \beta_i \rightarrow N_i$ without changing the units of $\eta$ ($1/\text{votes}$).

Hence we arrive at $\eta = \frac{c}{K_i}$, where $c$ is any constant that can be selected to tune the rate at which the variance approaches $0$.  In this study, we prefer $\eta = \frac{10}{K_i}$.

As a sanity check, one can verify that this form for $\eta$ satisfies properties $(1)-(3)$.

\section{Derivation of Eq.~\eqref{EQ3}}
\begin{align*}
\Pr(\alpha^*_i > T_i \mid X_i) &= 1 - \Pr(\alpha^*_i \leq T_i \mid X_i), \text{ where}\\
\Pr(\alpha^*_i \leq T_i \mid X_i) &= \int_0^1 \Pr(\alpha^*_i \leq T_i \mid p_i)\cdot \Pr(p_i \mid X_i) \mathrm{d}p_i\\
\end{align*}
Using the binomial and Beta distributions, we get
\begin{align*}
&= \int_0^1 \sum\limits_{v=0}^{V_i} {K_i \choose v} p_i^{v} {(1-p_i)}^{K_i-v} \cdot \operatorname{Beta}(p_i; 1+\frac{\alpha_i}{K_i},1+\frac{\beta_i}{K_i}) \mathrm{d}p_i,\\
&= \sum\limits_{v=0}^{V_i} \underbrace{\int_0^1 {K_i \choose v} p_i^{v} {(1-p_i)}^{K_i-v} \cdot \operatorname{Beta}(p_i; 1+\frac{\alpha_i}{K_i},1+\frac{\beta_i}{K_i}) \mathrm{d}p_i}_{= \Pr(\nu_i = v)\text{ for } \nu_i\ \sim\ \operatorname{Beta-Binomial}(1+\frac{\alpha_i}{K_i},1+\frac{\beta_i}{K_i}, K_i)}.
\end{align*}

Hence, we have
\begin{align*}
\Pr(\alpha^*_1 > T_1 \mid X_1) &= 1 - \sum\limits_{v=0}^{V_i} \Pr(\nu_i = v)\\
&= 1 - \Pr(\nu_i \leq V_i).
\end{align*}

\section{Derivation of Eq.~\eqref{EQ5}}
Assuming independence of votes across proposals,
\begin{align*}
\Pr\left(\bigwedge_{j \neq i} (\alpha^*_i > \alpha^*_j) \mid C\right) &=\prod_{j \neq i} \Pr(\alpha^*_i > \alpha^*_j \mid X_i, X_j).
\end{align*}
(If, instead, votes in $X_i$ were negatively correlated with votes in $X_j$, and $X_i$ had more support, then the equality ($=$) above would be replaced by an inequality ($>$).)

Further expanding the expression, we get
\begin{align*}
&= \prod_{j \neq i} \left( 1 - \Pr(\alpha^*_i \leq \alpha^*_j \mid X_i, X_j) \right)\\
&= \prod_{j \neq i} \left( 1 - \sum_{k=\lceil\delta_{ij}\rceil}^{K_j} \underbrace{\Pr(\alpha^*_j = \alpha_j + k \mid X_j)}_{= \Pr(\nu_j = k)} \cdot \underbrace{\Pr(\alpha^*_i \leq \alpha_j + k \mid X_i)}_{={CDF}_{\nu_i}(\alpha_j + k)} \right).
\end{align*}

\section{Derivation of Eq.~\eqref{EQ6}}
The properties we desire for $\lambda(t)$ are:
\begin{description}
\item[$(1)$] as $K_i \rightarrow 0$, the rate of incoming votes should drop exponentially and should be independent of $N_i$,
\item[$(2)$] as $t \rightarrow \infty$, the rate of incoming votes should drop exponentially.
\end{description}
Setting $\lambda (t) = \lambda_0 {(\gamma)}^{\frac{t}{K_i/N_i}}$ for $0 < \gamma < 1$ satisfies these constraints while still allowing for the exact rate to be tunable.

\section{Derivation of Eq.~\eqref{EQ7}}
Using $\lambda (t) = \lambda_0 {(\gamma)}^{\frac{t}{K_i/N_i}}$, we have
\begin{align*}
\lim_{t \rightarrow \infty} \Lambda (t) &=\int _{0}^{\infty}\lambda (t)\mathrm{d}t\\
&= \int _{0}^{\infty} \lambda_0 {(\gamma)}^{\frac{t}{K_i/N_i}} \mathrm{d}t\\
&= \lambda_0 \frac{K_i/N_i}{\ln{1/\gamma}}.
\end{align*}
Thus,
\begin{align*}
\lim_{t\rightarrow \infty}\Pr\left(N(t) = n\right) = \frac{(\lambda_0 \cdot K_i/N_i)^n}{n!(\ln{1/\gamma})^n}e^{-\lambda_0 \frac{K_i/N_i}{\ln{1/\gamma}}},
\end{align*}
and
\begin{align*}
\lim_{t\rightarrow \infty}\Pr\left(N(t)\leq \hat{K_i}\right) & = \sum_{n=0}^{\hat{K_i}} \frac{(\lambda_0 \cdot K_i/N_i)^n}{n!(\ln{1/\gamma})^n}e^{-\lambda_0 \frac{K_i/N_i}{\ln{1/\gamma}}}.
\end{align*}

%\begin{acks}
%\end{acks}

% Bibliography
\bibliographystyle{ACM-Reference-Format}
\bibliography{refs}

\end{document}